---
title: "Text Classification with R"
author: "Kalle Myllymaa"
date: "23 July 2018"
output:
  html_document:
    toc: true
    code_folding: "show"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Predicting the tone of a review

binary classification

# Exploratory data analysis

## Loading data

```{r data, cache = TRUE}
library(keras)
imdb <- dataset_imdb(num_words = 10000)
train_data <- imdb$train$x
train_labels <- imdb$train$y
test_data <- imdb$test$x
test_labels <- imdb$test$y
```

## Taking a look into the data
Training data (similarily to test data) consists of rows of vectors filled with ids for words.

```{r describe_train_data}
str(train_data[[1]])
```

Labels indicate if the review is considered as positive (1) or negative (0).

```{r describe_train_label}
train_labels[[1]]
```

```{r maxium, include=FALSE}
max(sapply(train_data, max))
```

Ids can be translated to words using a list mapping ids to string-presentations. For better performance, only 10 000 most common words are used and rest replaced with question marks.

```{r example_review}
library(keras)
word_index <- dataset_imdb_word_index()  
reverse_word_index <- names(word_index)
names(reverse_word_index) <- word_index

# Decodes the review. Note that the indices are offset by 3 because 0, 1, and 
# 2 are reserved indices for "padding," "start of sequence," and "unknown."
decoded_review <- sapply(train_data[[1]], function(index) {
  word <- if (index >= 3) reverse_word_index[[as.character(index - 3)]]
  if (!is.null(word)) word else "?"
})
cat(decoded_review)
```

# Data wrangling

As the data is presented in a vector of ids, it needs to be translated into a tensor.

```{r }
vectorize_sequences <- function(sequences, dimension = 10000) {
  # Creates an all-zero matrix of shape (length(sequences), dimension)
  results <- matrix(0, nrow = length(sequences), ncol = dimension) 
  for (i in 1:length(sequences))
    # Sets specific indices of results[i] to 1s
    results[i, sequences[[i]]] <- 1 
  results
}
```

```{r data_conversion}
x_train <- vectorize_sequences(train_data)
x_test <- vectorize_sequences(test_data)
```

```{r label_conversion}
y_train <- as.numeric(train_labels)
y_test <- as.numeric(test_labels)
```

# Creating a data model

```{r model_creation}
library(keras)

model <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>% 
  layer_dense(units = 16, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")
```

```{r model_compilation}
model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)
```

# Model validation

```{r validation_dataset_creation}
val_indices <- sample(1:25000, 10000, replace=F)

x_val <- x_train[val_indices,]
partial_x_train <- x_train[-val_indices,]

y_val <- y_train[val_indices]
partial_y_train <- y_train[-val_indices]
```

```{r validation}
val_history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 10,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)
```

```{r plotting_validation}
plot(val_history)
```

```{r predictions}
model %>% predict(x_test[1:10,])
```